---
title: "Life Insurance Assigment 2 Group 5"
output:
  pdf_document: default
  word_document: default
---
```{r install}
library(ggplot2)
library(gridExtra)
#install.packages("tinytex")
#tinytex::install_tinytex()
```

We imported the data as follows: Save a .txt file -> remove the title line of the text file -> Import Dataset (in RStudio) -> From Text (base) -> Seperator by Whitespaces -> use DATA object since file is named DATA.txt

Our data is for Luxembourg. It includes missing values, therefore affecting the plots and calculations, and the data is for years 1960-2022.

```{r setup, include=FALSE}
Force_of_mortality_a = function(x, t, i, j){
  subset_table = subset(DATA, DATA$Year == t)
  return(-log(1-subset_table$qx[i:j]))
}

Force_of_mortality_b = function(x, t, i, j) {
  subset_table = subset(DATA, DATA$Age == x)
  return(-log(1-subset_table$qx[i:j]))
}
```

Part A below

```{r cars}
# assume constant force of mortality 
colors_years = c("1960" = 'black', "1970"='purple', "1980"='red', "1990"='blue', "2000"='green', "2010"='yellow', "2020"='orange')

for (x in c(1960, 1970, 1980, 1990, 2000, 2010, 2020)) {
  forces_of_mortality = Force_of_mortality_a(0, x, 1, 109)
  print(forces_of_mortality)
  if (x == 1960) {
    plot(1:length(forces_of_mortality), log(forces_of_mortality), type = "l", xlab = "Year", col='black', ylab = "Force of Mortality", main= "Force of Mortality vs. Year for Ages 0-110", ylim=c(-10, 0))
  }
  else {
    lines(1:length(forces_of_mortality), log(forces_of_mortality), type = "l", pch=18, col = colors_years[as.character(x)])
  }
}
legend("topleft", legend=c("1960","1970"," 1980", "1990", "2000", 
                            "2010", "2020"), 
       col= c("black","purple","red", "blue", "green", 'yellow', 'orange'), lty=1, cex=0.5)

```
```{r description1}
# for different selected time periods t, once again, in order to not encounter issues relating to the different 
# future expectations for different (x) year old, a cohort table was used as this tracked the lives of people of differing ages in time.
```

Part B Below

```{r pressure2, echo=FALSE}
#b μx,t the force of mortality of an (x) year old in year t, versus t, 
# for different selected time periods t

# assume constant force of mortality 
colors_ages = c("0" = 'black', "10"='purple', "20"='red', "30"='blue', "40"='green', "50"='yellow', "60"='orange', "70"='grey', "80"='cyan')

for (x in c(0, 10, 20, 30, 40, 50, 60, 70, 80)) {
  forces_of_mortality = Force_of_mortality_b(x, 1960, 1, 2022-1960)
  if (x == 0) {
    plot(1960:2021, log(forces_of_mortality), type = "l", xlab = "Year", col='black', ylab = "Force of Mortality", main= "Force of Mortality vs. Year for Ages 0-80", ylim=c(-10, 0))
  }
  else {
    lines(1960:2021, log(forces_of_mortality), type = "l", pch=18, col = colors_ages[as.character(x)])
  }
}
legend("topright", legend=c("0-years-old","10-years-old"," 20-years-old", "30-years-old", "40-years-old", 
                            "50-years-old", "60-years-old", "70-years-old", "80-years-old"), 
       col= c("black","purple","red", "blue", "green", 'yellow', 'orange', 'grey', 
              'cyan'), lty=1, cex=0.4)
```

Part C Below

```{r description2}
# part c
survival_function<-function(x,t){
  px<-1-DATA$qx[DATA$Year==t]
  if(x==0&t>=min(DATA$Year)&t<=max(DATA$Year)){
    survival<-1
    print(survival)
  }
  if(x>=110&t>=min(DATA$Year)&t<=max(DATA$Year)){
    survival<-0
    print(survival)
  }
  if(x>0&x<110&t>=min(DATA$Year)&t<=max(DATA$Year)){
    survival<-prod(px[1:x])
    print(survival)
  }
  if(x<0|t<min(DATA$Year)&t>max(DATA$Year)){
    sprintf("Impossible to calculate the survival function with t=%d and x=%d",t,x)
  }
}

colors_years = c("1960" = 'black', "1970"='purple', "1980"='red', "1990"='blue', "2000"='green', "2010"='yellow', "2020"='orange')
for (t in  c(1960, 1970, 1980, 1990, 2000, 2010, 2020)) {
  px<-1-DATA$qx[DATA$Year==t]
  kpx <- cumprod(px[(0 + 1):(108 + 1)])
  if (t == 1960) {
    plot(0:108, sqrt(kpx), type='l', xlab = "Age", col='black', ylab = "Survival Probability", main= "Survival Probabilities vs. Age in Various Periods 1960-2020", ylim=c(0,1))
  }
  else {
    lines(0:108, sqrt(kpx), type='l', col=colors_years[as.character(t)])
  }
}
legend("bottomleft", legend=c("1960","1970"," 1980", "1990", "2000", "2010", "2020"), 
       col= c("black","purple","red", "blue", "green", 'yellow', 'orange'), lty=1, cex=1)

```
```{r pressure31, echo=FALSE}
# GOAL: compare the forces of mortality of x year old in 1970 vs 1990 vs 2010
```

With this, the objective is to determine weather the progression of time and technological improvements, 
cause the forces of mortality to decrease for all ages. As shown in the graphs, and as could be expected by looking at real world exmaples,
as time progress, i.e. 1970 -> 1990 -> 2010, medical advancements combined with a myriad of other factors cause a decrease in the forces of mortality the board.


Part D below


```{r pressure333, echo=FALSE}
colors_years = c("1960" = 'black', "1970"='purple', "1980"='red', "1990"='blue', "2000"='green', "2010"='yellow', "2020"='orange')
for (t in  c(1960, 1970, 1980, 1990, 2000, 2010, 2020)) {
  px<-1-DATA$qx[DATA$Year==t]
  kpx <- cumprod(px[(0 + 1):(20 + 1)])
  if (t == 1960) {
    plot(0:20, kpx, type='l', xlab = "Age", col='black', ylab = "Survival Probability", main= "Survival Probabilities vs. Age of a Newborn Reaching Adulthood (Age 21)", ylim=c(0.8,1))
  }
  else {
    lines(0:20, kpx, type='l', col=colors_years[as.character(t)])
  }
}
legend("bottomleft", legend=c("1960","1970"," 1980", "1990", "2000", "2010", "2020"), 
       col= c("black","purple","red", "blue", "green", 'yellow', 'orange'), lty=1, cex=1)
```
